---
title: "Walinchus Astronomy"
author: "David H. Montgomery"
date: "10/19/2020"
output: html_document
---

```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(scales)
library(ggforce)
library(ggrepel)
```

## Load our data

The first step is to always load your data in, and clean up any messy bits

```{r loaddata}
# Load the data
space_data <- read_csv("https://raw.githubusercontent.com/dhmontgomery/walinchus-astronomy/master/space_data.csv") %>%
	janitor::clean_names() %>% # Run a function to clean up the column names
	mutate(across(-object, as.numeric)) %>% # Convert everything to numeric
	filter(!is.na(distance_from_sun_million_km)) %>% # Remove missing data
	mutate(object = str_remove_all(object, "\\(an asteroid\\)")) # Remove "(an asteroid)" from Ceres' name
```

## Create a quick graph: distance from the Sun

The first thing I always like to do is make a quick, ugly chart to see what the data looks like.

```{r distance_plot, message = FALSE}
# Format the data
space_data %>% # Start with our data	
	# Create the graph
	ggplot(aes(x = fct_reorder(object, distance_from_sun_million_km), # Set X as the object names, ordered by distance
			   y = distance_from_sun_million_km)) + # Set Y as the distance
	geom_col() + # Make it a column graph
	scale_y_continuous(labels = comma) + # Format the y-axis text
	# Set chart labels
	labs(title = "Distance from the Sun", # Title
		 y = "Millions of kilometers", # Y-axis title
		 x = "Celestial body", # X-axis title
		 caption = "Graph by David H. Montgomery") # Caption
ggsave("distance_from_the_sun.png", width = 8, height = 6) # Save to disk
```

## Make our graph prettier

Once I've done that, I customioze the chart to make it look nicer — something I might want to publish.

```{r distance_plot_pretty, message = FALSE}
# Format the data
space_data %>% # Start with our data
	# Create the graph
	ggplot(aes(x = fct_reorder(object, distance_from_sun_million_km, .desc = TRUE), # Set X as the object names, ordered by distance
			   y = distance_from_sun_million_km)) + # Set Y as the distance
	geom_col() + # Make it a column graph
	# Add direct labels 
	geom_text(aes(label = comma(distance_from_sun_million_km, accuracy = .1), # Set the label text
				  # Set a different position for large and small values
				  hjust = case_when(distance_from_sun_million_km < 500 ~ -.1, TRUE ~ 1.1), 
				  # Set a different color for large and small values
				  color = case_when(distance_from_sun_million_km < 500 ~ "black", TRUE ~ "white"))) +
	scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.05))) + # Format scale
	scale_color_identity() + # Tell R to treat "white" and "black" as actual colors
	coord_flip() + # Rotate the chart 90º
	# Set chart labels
	labs(title = "Distance from the Sun (millions of kilometers)",
		 y = "Millions of kilometers",
		 x = "Celestial body",
		 caption = "Graph by David H. Montgomery") +
	theme_minimal() + # Set a theme
	# Customize theme
	theme(axis.title = element_blank(), # Hide axis titles
		  axis.text.x = element_blank(), # Hide x-axis labels
		  # Hide vertical griddlines
		  panel.grid.major.y = element_blank(),
		  panel.grid.minor.y = element_blank())
ggsave("distance_from_the_sun_pretty.png", width = 8, height = 6) # Save to disk
```

## Make a complicated graph, showing distance as orbits

I often experiment with different ways of visualizing the data. The below chart takes the same data as above — each celestial object's distance from the Sun — and portrays it with approximations of orbits, a possibly more intuitive way to understand the data.

```{r radius, message = FALSE, warning = FALSE}
space_data %>% # Start with our data
	ggplot() + # Start graph
	geom_circle(aes(x0 = 0, y0 = 0, # Set our origin to (0, 0) 
					r = distance_from_sun_million_km), # Use distance as our radius
				size = .2, color = "grey90") + # Set size and color of lines
	geom_point(aes(x = 0, y = distance_from_sun_million_km, # Add points
				   size = diameter_km)) + # Govern dot size by diameter
	# Add blank labels as a background (hacky workaround to make this look better)
	geom_label(data = . %>% filter(distance_from_sun_million_km > 500), aes(x = 0, y = distance_from_sun_million_km, label = object), hjust = 0, nudge_x = 350, color = NA, alpha = .5) + 
	# Add text labels
	geom_text(data = . %>% filter(distance_from_sun_million_km > 500), # Filter out inner planets
			  aes(x = 0, # x-axis position to 0
			  	y = distance_from_sun_million_km, # y-axis position by distance
			  	label = object), # Label from object name
			  hjust = 0, nudge_x = 350) + # Set the position of the text relative to the dots
	scale_size_continuous(labels = comma, breaks = c(1000, 10000, 100000)) + # Set legend display
	coord_fixed() + # Set a fixed coordinate scale
	theme_void() + # Hide grid, axes, etc.
	# Label chart
	labs(title = "Distance from the Sun",
		 caption = "Graph by David H. Montgomery",
		 size = "Diameter (km)")
ggsave("distance_from_the_sun_circles.png", width = 8, height = 6) # Save to disk
```

You can't legibly show the inner planets and outer planets on the same scale. What are some solutions to this?

## Plot min and max temperatures

A key part of data analysis is to understand the relationship between two different variables. Here I plot each celestial body's minimum and maximum temperature. Here we can see all the planets fall below freezing sometimes — except Venus, by far the hottest planet in the solar system. Why is Venus so hot?

Meanwhile, the rest of the planets can all get pretty cold, but they vary wildly in terms of how hot they can get. Some planets are cold all the time, Mercury goes from super cold to super hot, and Earth and Mars max out at milder temperatures.

```{r}
space_data %>% # Start with our data
	# Make the chart
	ggplot(aes(x = temperature_min_c, # x = min temperature
			   y = temperature_max_c)) + # y = max temperature
	geom_point() + # Make it a scatterplot
	geom_text_repel(aes(label = object)) + # Add labels, repelled away if they overlap
	theme_minimal() + # Set our theme
	coord_fixed(clip = "off") + # Fix coordinates, turn off clipping at the edge of the graph
	# Label chart
	labs(title = "Minimum and maximum temperatures",
		 caption = "Graph by David H. Montgomery",
		 x = "Minimum temperature (Celsius)",
		 y = "Maximum temperature (Celsius)")
ggsave("temperature.png", width = 7, height = 7) # Save to disk
```

## Learning more

Various R tutorials I've written:

- [R data analysis & visualization for beginners](https://github.com/dhmontgomery/r-data-for-beginners)
- [Gathering and cleaning data in R](https://github.com/dhmontgomery/nicar19/blob/master/r3.md)
- [Making graphics with R](https://github.com/dhmontgomery/nicar20/blob/master/ggplot-graphics/readme.md)

Hadley Wickham and Garrett Grolemund's excellent free ebook, [R for Data Science](https://r4ds.had.co.nz/)

The video game [Kerbal Space Program](https://www.kerbalspaceprogram.com/)

CGP Grey's video, ["Which Planet is Closest"](https://www.youtube.com/watch?v=SumDHcnCRuU):

<iframe width="560" height="315" src="https://www.youtube.com/embed/SumDHcnCRuU" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>